---
title: 'Part1: Simulation Exercise'
author: "Haonan"
date: "October 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This is a simulation practice which we would like to compare the simulation means and variances with the theoratical means and variances of a certain distribution. The distribution is an exponential distribution with lamda equals 0.2, 40 simulations will make up a sample and totally there are 1000 samples. This simulation will produce mean and variance plots and verify the assumption that the mean and the variance distribution is approximately normal.

# Set Up the Simulation

We use the code as follows to set up the simulation results:

```{r }
setwd("C:/Users/Haonan/Desktop/Study/R/statistical inferring project")

library(ggplot2)

random <- data.frame()
for (i in 1:1000){
  a <- rexp(40,.2)
  meanr <- mean(a)
  stdr <- sd(a)
  random[i,1] <- meanr
  random[i,2] <- stdr
}
colnames(random) <- c("Mean","Std_Dev")

head(random)
```

#Comparison

## Sample Means

The theoretical mean of exponential distributions is 1/lamba, which, in this case, was 1/0.2 = 5
Firstly, it is reasonable to conduct a t test to see whether there's a significant difference between the sample mean and the theoretical mean
```{r}

t.test(random$Mean, mu = 5)
```

From the result we could clearly identify that the p-value is very large. This means there's no significance between the sample mean and the theoretical mean, they're equal. 

Then we graph it, add a vertical line which indicates the theoretical mean = 5, and finally add a normal line as follows:
```{r}
gg1 <- ggplot(random, aes(Mean)) + geom_histogram(alpha = .3, fill = 3, binwidth = .1, aes(y = ..density..))+ stat_function(fun=dnorm, size = 3, color = "red", args=list(mean=mean(random$Mean), sd=sd(random$Mean))) + geom_vline(xintercept = 5, size = 2)+labs(x = "Mean", y = "Density", title = "Sample Mean Density Distribution")
gg1
```

From the plot, the sample mean is almost normally distributed with mean = 5. Compared with the red line, the sample mean is approximately normal

## Sample Variance

The theoretical variance of exponential distributions is 1/lamba^2, which, in this case, was 1/0.2^2 = 25
Firstly, it is reasonable to conduct a t test to see whether there's a significant difference between the sample mean and the theoretical mean

```{r}

with(random, t.test(Std_Dev^2, mu = 1/.04))
```

From the result we could clearly identify that the p-value is quite large. This means there's no significance between the sample variance and the theoretical variance, they're equal

Then we graph it, add a vertical line which indicates the theoretical variance = 25, and finally add a normal line as follows:
```{r}
gg2 <- ggplot(random, aes(Std_Dev^2)) + geom_histogram(alpha = .3, fill = 4, binwidth = 2, aes(y = ..density..))+ stat_function(fun=dnorm, size = 3, color = "red", args=list(mean=mean(random$Std_Dev^2), sd=sd(random$Std_Dev^2)))+ geom_vline(xintercept = 25, size = 2) + labs(x = "Variance", y = "Density", title = "Sample Variance Density Distribution")
gg2
```

From the plot, the sample variance seems like normally distributed with mean = 25, but still, a little skewed to the right. Compared with the red line,  the sample variance distribution is approximately normal

# Conclustsion

The simulation exercise has produced the desired results that the sample mean and the sample variance is nearly normal




